\setstretch{1.05} SUPPORTING INFORMATION:\\ Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project (\url{nmrlipids.blogspot.fi})}

### 2015-11-24 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Minor changes in the proof stage

Since Poger lipids come from the GROMOS force field, it is important to note that GROMOS uses the RF scheme for computing electrostatics (this is the method used for the 
force field parameterization). Using setup i) based on RF, we were able to reproduce the results (i.e. area per lipid value of 0.63~nm$^2$) from the original work only with 
GROMACS versions 4.0.X and earlier (the original authors~\cite{poger10} used GROMACS version 3.3.3). When switching to versions 4.5.X and above, the area per lipid dropped to below 0.58~nm$^2$. 
The GROMACS developers were contacted and a redmine issue opened (\url{http://redmine.gromacs.org/issues/1400}). Outcome: the difference comes from the new Trotter decomposition 
introduced in versions 4.5.X. A fix has been introduced in version 4.6.6 that allows a recovery of an area per lipid value of 0.615~nm$^2$. The results in terms of area per lipid using the different 
GROMACS versions are available at Ref.~\citenum{pogerFILESrf2}.

@@

\subsection{Botan-CHARMM36-UA}
This model was similar to the Lee-CHARMM36-UA model with a few  differences.
A hydrated bilayer consisting of 128 DLPC lipids and 3840 water
molecules was described by a model derived from the CHARMM27-UA model by H\'enin et al.~\cite{henin08}.

>>>>

Since Poger lipids come from the GROMOS force field, it is important to note that GROMOS uses the RF scheme for computing electrostatics (this is the method used for the 
force field parameterization). Using setup i) based on RF, we were able to reproduce the results (i.e. area per lipid value of 0.63~nm$^2$) from the original work only with 
GROMACS versions 4.0.X and earlier (the original authors~\cite{poger10} used GROMACS version 3.3.3). When switching to versions 4.5.X and above, the area per lipid dropped below 0.58~nm$^2$. 
The GROMACS developers were contacted and a redmine issue opened (\url{http://redmine.gromacs.org/issues/1400}). Outcome: the difference comes from the new Trotter decomposition 
introduced in versions 4.5.X. A fix has been introduced in version 4.6.6 that allows a recovery of an area per lipid value of 0.615~nm$^2$. The results in terms of area per lipid using the different 
GROMACS versions are available at Ref.~\citenum{pogerFILESrf2}.

@@

\subsection{Botan-CHARMM36-UA}
This model was similar to the Lee-CHARMM36-UA model with a few  differences~\cite{botan15}.
A hydrated bilayer consisting of 128 DLPC lipids and 3840 water
molecules was described by a model derived from the CHARMM27-UA model by H\'enin et al.~\cite{henin08}.
